Objectification in Action: Self- and Other-Objectification in Same-gender Interactions

### 2020-09-09 Final_Manuscript.tex
re-knit paper

\abstract{
Empirical evidence has only found links between objectification,
self-objectification, and negative outcomes for women within
interpersonal interactions between male-female pairs. The purpose of the
present study was to extend past research and consider the relationships
between such valuable phenomena and their effects on authenticity within
interactions between female pairs. Women were brought into the
laboratory and interacted in same-sex dyads. Dyadic analysis was
utilized to detect whether partners' objectification of each other
affected state self-objectification, and the resulting feelings of
comfort and authenticity during the interaction. After the interaction,
participants completed a questionnaire which measured many constructs
including cognitive performance, career aspirations, and relationship
agency. Results revealed no significant relationship between
self-objectification and authenticity. Further, although there were
significantly negative effects on career aspirations and relationship
agency resulting from a lack of relationship authenticity, there was no
evidence that this is due to feelings of sexual objectification. The
significant partner effect of objectification on actor
self-objectification suggests that women being objectified by other
women still results in feelings of self-objectification, and such
research has powerful implications for the ways that women interact in
both sexual and non-sexual settings. AUTHOR NOTE: mention Clark trigram
coders Hannah et al.

@@

Roberts and Fredrickson (1997) Objectification Theory, suggests that, in
addition to being steeped in a culture that objectifies women, women are
objectified in actual interpersonal encounters. The negative effects of
this interpersonal objectification might be the strongest when it is a
man doing the objectifying (Gervais, Vescio, \& Allen, 2011), but what
are the consequences when a woman objectifies another woman in an
interaction? Psychological researchers studying the sexual
objectification of women, have in recent years explored the
interpersonal process of the objectification, finding evidence for
self-objectification in the target as a proximal consequence (Garcia,
Earnshaw, \& Quinn, 2016, Strelan and Pagoudis (2018), Riemer, Sáez,
Brock, and Gervais (2020)). For example, in a face-to-face mixed-gender
dyadic study, Garcia et al. (2016) found that men's reported
objectification of their female interaction partner was associated with
increased self-objectification reported by their partner. They did not
include same-gender, woman-woman interacting pairs. Gervais, Sáez,
Riemer, and Klein (2020) have reviewed the literature on interpersonal
self-objectification and has organized the process in a theoretical
model called the Social Interaction Model of Objectification (SIMO), but
this model is currently only relevant for mixed-gender encounters.
Although there is ample evidence that women objectify other women
(Loughnan et al., 2015a; Puvia \& Vaes, 2013), there are currently no
studies investigating the process of interpersonal objectification in
same-gender interactions. Does objectification by a female interaction
partner have the same downstream negative consequences as being
objectified by a male interaction partner? The current dyadic study
addresses this omission using interacting pairs where both members
identify as women.

@@

Harsey and Zurbriggen (2020) found that self-objectification was related
to the objectification of women to a similar degree for men and women
participants (although men scored higher than women on a measure of
belief that objectification of women is natural for men). Additionally,
Strelan and Hargreaves (2005) found that the more women self-objectify,
the more they objectify other women. Could the process work in the other
direction as well, that is, could objectification by a woman produce
self-objectification, in a similar manner to mixed-gender interactions?

{[}Asha comment on next paragraph: Hill and Fischer (2008) assessed
women's experiences of objectification from men separately from women's
experiences of objectification from women, and found that women may be
socialized not only to see themselves as objects, but perhaps to see
other women as objects as well, although this process was not comfirmed
to occur during an actual interpersonal interaction

Could this sentence be written like:

Though, when Hill and Fischer(2008) assessed women's experiences of
objectification from men independently from that of women, they found
that women are likely to be socialized to not only to see themselves s
objects, but potentially see other women as objects as well. This
process ,however, was not confirmed to occur during an interpersonal
interaction.{]}

Past research has focused on men's objectifying behaviors, with little
attention paid to the role that women may also play in objectifying
other women. Hill and Fischer (2008) assessed women's experiences of
objectification from men separately from women's experiences of
objectification from women and found that women may be socialized not
only to see themselves as objects but perhaps to see other women as
objects as well, although this process was not comfirmed to occur during
an actual interpersonal interaction. Loughnan et al. (2015a) found that
women objectify other women to a greater extent than they objectify men.
Women have also been found to objectify (dehumanized) sexualized targets
presented as images (Puvia \& Vaes, 2013), but very little is known
about interpersonal sexual objectification among women and what the
proximal effects of this objectification might be. We do know that
objectification has more adverse consequences for women than men
(Gervais et al., 2011; Moradi \& Huang, 2008; Saguy, Quinn, F Dovidio,
\& Pratto, 2010), however, if women objectify other women, perhaps even
to a stronger degree than they objectify men, this higher level of
objectification might be a detrimental combination.

@@

Studies have shown that within social encounters women are gazed at more
than men (Briton \& Hall, 1995), and often times feel \enquote{looked
at} within interpersonal interactions (Argyle \& Williams, 1969;
Miles-McLean et al., 2015), and will more than likely internalize the
objectifying gaze on their physical self (Puvia \& Vaes, 2013; Young,
1979). Moreover, perhaps the most adverse effect is that it socializes
girls and women to treat themselves as objects to be looked at and
evaluated (Bartky, 1990; Berger, Cohen, \& Zelditch Jr, 1972; B. L.
Fredrickson et al., 1998). Gervais et al. (2020) recently reviewed the
research on interpersonal objectification and organized our current
theoretical understanding of this process in the Social Interaction
Model of Objectification (SIMO). What is clear from Gervais et al.
(2020) is that we know little about women-on-women interpersonal
objectification. Could there also be negative consequences of this type
of interpersonal objectification? Perhaps women could be amplifying each
others' state self-objectification as they get ready to go out for the
night {[}EDIT??{]}, \emph{before} the theoretical encounters with men
organized in the SIMO occur. In this way, women's rituals of getting
ready together might place even women with low TSO in a higher SSO
group. {[}Kat: This has me thinking about the difference between
objectification in a romantic setting (which was the setting created in
the initial experiment) vs objectification in a platonic setting/between
friends. Might be useful to discuss this difference, or just have a
sentence here that clarifies objectification can occur within non-sexual
settings{]} {[}Asha: I think this is a great way to illustrate the
potential experience but when I read it I was kind of like
\enquote{wait, when did we talk about women going out?}{]}

Interpersonal objectification can also occur within the context of
romantic relationships, in addition to interactions with strangers.
Among heterosexual male and female college students,
self-objectification has been found to be positively associated with the
extent to which they objectify their romantic partners (Zurbriggen,
Ramsey, \& Jaworski, 2011). Researchers have also studied interpersonal
objectification and self-objectification in romantic relationships
(Strelan \& Pagoudis, 2018). Evidence is mixed, suggesting that women
can sometimes have positive outcomes from appearance valuation within
the context of an established romantic relationship. There is also
evidence that they experience less SSO after appearance comments within
romantic relationships (Meltzer, 2020), but less is known about first
dates. An important caveat to this literature is that it is only about
heterosexual relationships. A recent study has investigated the effect
of self-objectification on the reduction of relationship building skills
in general (including same-sex friendships) {[}CITE PWQ paper I
reviewed, Yoder editor{]}. Both romantic relationships and stranger
encounters have been studied in the context of objectification, but
neither of these literatures have investigated specifically same-gender
relationships.

There is evidence that the experience of state self-objectification in
mixed-gender contexts (stranger and romantic) has negative consequences
for women, but what about in the context of interactions with other
women? Is the self-objectification experienced in mixed-gender
interactions associated with the same negative process as that
experienced in same-gender interactions? On the one hand, there is ample
research demonstrating that the male gaze has a particularly detrimental
effect would suggest no (Gay \& Castano, 2010; Gervais et al., 2011;
Saguy et al., 2010, Calogero (2004); Yilmaz \& Bozo, 2019), but on the
other hand, women do objectify other women (Harsey \& Zurbriggen, 2020;
Loughnan et al., 2015a; Puvia \& Vaes, 2013). When they do it, does it
lead to self-objectification in the same way men's objectification of
women does (Garcia et al., 2016)? If so, does the self-objectification
experienced in these same-gender interactions have the same negative
consequences for authenticity in that interaction?

@@

There is some evidence that reduced authenticity is a consequence of
self-objectification in the moment {[}Garcia et al. (2016); again CITE
PWQ paper I reviewed, Yoder editor{]}. This link has been justified by
the literature on stigmatized-stigmatizer interactions (M. R. Hebl \&
Dovidio, 2005; Richeson \& Shelton, 2003; Shelton, Richeson, \&
Salvatore, 2005), viewing the experience of being objectified in an
interaction as an identity threat situation (Miles-McLean et al., 2015;
Nadal \& Haynes, 2012). Further, empirical evidence reveals that
objectification manifests through inauthenticity in romantic
relationships (Brunell et al., 2010), adverse attitudes in regard to
career aspirations, and a decrease in concentration and impairment in
female cognitive performance (Kahalon, Shnabel, \& Becker, 2018; D. M.
Quinn, Chaudoir, \& Kallen, 2011). When a woman is objectified by a man,
and subsequently experiences self-objectification, the intergroup nature
of this encounter might trigger identity threat. Thus, perhaps there are
fewer negative consequences when a woman is objectified by another
woman, specifically, woman-woman interpersonal objectification processes
might diverge from mix-gender interpersonal objectification processes
precisely because they are not intergroup interactions (at least with
respect to gender identity). A woman objectified by another woman may
not be having the same negative consequences that cascade from
situations that trigger group-based identity threat (Deaux \& Major,
1987; Dovidio, Hebl, Richeson, \& Shelton, 2006; M. R. Hebl \& Dovidio,
2005), but feeling like a body, rather than a full human, in any
interaction (intergroup or ingroup) may be enough to reduce women's
feelings of authenticity and social competence, regardless of the gender
of the objectifyer {[}Tolman, Impett, Tracy, and Michael (2006); PWQ
paper I reviewed, Yoder editor{]}.

@@

In the current study, we sought to examine what occurs during an
interaction in which one or both partners are objectifying each other,
similarly to Garcia et al. (2016), but between same-gender interpersonal
interactions among women. Moreover, the current study uses a
face-to-face interaction paradigm and dyadic data analysis techniques to
examine the effects for both women simultaneously. Although the
literature on woman-woman objectification is small, the results are
mixed, and does not cover interpersonal encourters, we expected to
replicate some of the results found in Garcia et al. (2016). Most
importantly, we predicted that being objectified by one's interaction
partner would lead to self-objectification. We also expected, in line
with Garcia et al. (2016), that TSO would moderate this relationship,
amplifying the positive assocaition between being objectified and SSO
for women higher in TSO. We hypothesized that SSO would, in turn, lead
to feelings of inauthenticity (although there is ample support against
this prediction from the identity threat within intergroup interactions
literature; Rollero, 2016). Further, we hypothesized that these feelings
of inauthencity would then be associated with reduced feelings of agency
in romantic relationships, reduced career aspiration, and reduced
cognitive performance. In summary, we expected to find a positive
relationship between other-objectification by one's partner and state
self-objectification. We also expected to find a negative relationship
between self-state objectification and interaction authenticity, and
that interaction authenticity will be positively related to cognitive
performance, relationship agency, and career aspirations.

Here we would like to note that Puvia and Vaes (2013) would
alternatively predict that women's tendency to self-objectify (TSO)
leads them to objectify other women (more precisely, to dehumanize a
sexualized woman). In addition, Strelan and Hargreaves (2005) and Harsey
and Zurbriggen (2020) would predict the TSO to other-objectification
link, further theorezing that this relationships is mediated by the
woman's own state self-objectification (SSO). This limited, but extant,
set of literature on woman-woman objectification points to a possible
alternative model from the model hypothesized in the current study.
Where appropriate, we report results considering the causal directions
implied by this alternative mediation model.

@@

\emph{All} participants were told that they were assigned to the
\enquote{College Relationships} condition. Self-objectification has been
found to occur after a mere relationship prime among women (Sanchez \&
Broccoli, 2008) because, in Westurn culture and beyond, women need to
look attractive to obtain and maintain successful relationship, thus,
the \enquote{College Relationships} condition may heighten
self-objectification and the evaluation of other women in a sexualized
way. The decision was made to ask even heterosexual women to judge their
fellow-woman partners for their potential as romantic partners. We felt
that this prompt would keep the study closest to a replication of the
previous Garcia et al. (2016) version of the study, and strengthen the
potential for objectification that would normally be low in the context
of a psychology laboratory. Recall that, past research has found that
women are indeed able to evaluate other women's potential as romantic
partners {[}i.e., their sexual attractiveness; CITE{]}---indeed, women
may be unfortunately quite used to thinking about their own
\enquote{datability,} and we suspect this this habitual thought pattern
will translate to their thoughts about other women.

@@

\subsection{Combined Samples}\label{combined-samples}

Data from two different, but demographically equivalent, samples were
combined to create the final analysis sample (\emph{N =} 64) used in
this study. In the measures section that follows we refer to them as
Sample 1 and Sample 2. Thirty-two previously unacquainted
self-identifying female-sex dyads (64 total participants) from two
liberal arts institutions in the Northeast of the United States
participated in this study. Sample 1 (\emph{N =} 24) is from a co-ed
liberal arts college in the northeast US and Sample 2 (\emph{N =} 40) is
from a women's liberal arts college in the northeast US. More
specifically, twelve of the pairs, which derived from Sample 1, were
students at a co-ed liberal arts college, while the remaining twenty
pairs who came from Sample 2 attended a women's liberal arts college.
Initially, data was collected from same-sex and mixed-sex dyads that
comprised of male and female gendered individuals. Sample 1 originally
consisted of twenty-two pairs, twelve men and thirty-two women.
Twenty-three pairs made up of forty-three women and one man, as well as
two participants who did not identify with either gender category,
formed Sample 2. For consistency, we limited participant data to same
sex female pairs at the two colleges.

Due to these similarities across samples in regard to correlation
patterns between significant variables within this study, the two
datasets were combined. These participants were mostly first-year
college students, with an average age of 18.85 (SD = 1.04). The sample
was 48.44\% White/European American, 9.38\% Black/African-American,
28.12\% Asian/Pacific Islander, 9.38\% Latinx, and 4.69\% mixed-race.
There were 8 White/White pairs and 4 same race racial minority pairs,
for a total of 12 same-race pairs. The remaining 20 were mixed race
pairs, of which 15 were White/racial minority pairings and 5 were
cross-racial minority group pairs. 64.06\% of the sample identified as
heterosexual, and 25\% identified as gay, lesbian or bisexual.

@@

To measure the participant's objectification of their partner in the
interaction, participants were asked a series of questions about the
frequency of thoughts in relation to multiple characteristics of their
partner (Garcia et al., 2016). Questions included aspects of their
partner's internal traits such as personality, friends, family, and
extracurricular interests, as well as external traits such as body,
appearance, clothing, and body parts. All questions were rated on a
scale from 1 (not at all) to 7 (constantly). Objectification was
measured by getting the difference between the average frequency of
thought about their partner's external traits (\(\alpha\) = 0.79 for
Sample 1, \(\alpha\) = 0.79 for Sample 2) and frequency of thought about
their partner's internal traits (\(\alpha\) = 0.79 for Sample 1,
\(\alpha\) = 0.76 for Sample 2). A positive score in this scale would
indicate that the participant thought about their partner's external
traits more than the partner's internal traits, and a negative score
would indicate the opposite.

@@

\begin{figure}
\includegraphics[width=400px]{2016_figure} \caption{Path Analysis Model from Garcia et al. (2016) study with distinguishable dyads.}\label{fig:2016figure}
\end{figure}

This study sought to replicate the results of Garcia et al. (2016)'s
study, done with male-female pairs, which used a dyadic path analysis to
detect whether partners' objectification of one another affected state
self-objectification (SSO). See Figure~\ref{fig:2016figure} for the
results of the analysis from this previous study. We investigated how
the central effects found in the previous study relate to interactions
between two women. Specifically, we are interested in testing the
relationship between state-other objectification and SSO, and how SSO in
turn, affects feelings of inauthenticity during the interaction. In
addition, we will also test if the effect of other-objectification in an
interaction on SSO is only present for those women who are high in trait
self-objectification, as in Garcia et al. (2016). Further, we will
investigate the relationships between experiencing interaction
inauthenticity and relationship agency, career aspirations, and
cognitive performance. {[}Asha: This section feels like a mixture of
explaining what the study was intended to investigate as wel as future
directions. Is that the point, or can this be organized better?{]}

While Garcia et al. (2016) used dyadic path analysis, the current study
used multilevel modeling for our dyadic analyses. Dyadic analyses for
distinguishable dyads (e.g., mixed-gender interacting pairs) is more
natural in Structural Equation Modeling (SEM) than it is for
indistinguishable dyads (e.g., same-gender interacting pairs) (Garcia,
Kenny, \& Ledermann, 2015; Ledermann \& Kenny, 2017). One reason for
this asymmetry is that, due to the arbitrary distinctions made between
\enquote{partner 1} and \enquote{partner 2} in indistinguishable dyads,
many estimates need to be fixed to be equal (i.e., paths, variances,
covariances, endogenous intercepts, and exogenous means) for
indistinguishable dyads but these equality constraints should not then
be considered in the degrees of freedom calculations for fit estimations
(Olsen \& Kenny, 2006). Further, Olsen and Kenny (2006) detail how a new
independence model and the corresponding fit measure should be
re-calculated for indistinguishable dyads models. The current study uses
dyadic multilevel modeling (MLM) to test all relationships and mediation
patterns. The online supplementary materials contains analysis using
SEM. See Ledermann and Kenny (2017) for a more complete discussion of
the considerations for using SEM versus MLM for dyadic analysis.

Testing the Garcia et al. (2016) model on the current, same-gender,
sample, involves using the Actor-Partner Independence Model (APIM)
approach for each outcome variable (i.e., endogenous variable in
Figure~\ref{fig:2016figure}). Thus, we ran five APIM's to test all the
hypothesized relationships. See Figure~\ref{fig:apim} for a basic APIM
model. The APIM includes effects due to one's own, as well as one's
partner's, predictor variables (\(X\)'s) on the one's own outcome
variable (\(Y\)). Unlike the original Garcia et al. (2016) study, our
study deals with indistinguishable dyads, meaning the designation of who
is designated as \enquote{actor} and who is designated as
\enquote{partner} is arbitrary. Recall that the indistinguishable nature
of the dyads in the current study led us to choose the MLM approach over
SEM. These analyses are considered exploratory, given the lack of prior
research theorizing about these linkages.

Before moving to the main analyses, we discuss statistical equivalence
tests that provide support for combining Sample 1 and Sample 2 in one
analysis sample. The online supplemental material contains the main
analyses separated by samples. All results presented below are from
models including sample as a control variable.

\subsubsection{Combining Samples}\label{combining-samples}

The correlations between study variables is similar across samples. The
reliabilities for the study scales were also equivalent. (Note that the
two samples were two small to conduct formal measurement equivalence
tests for scales.) There are no statistically significant differences
between samples in demographics including age, STATS, and ethnicity,
STATS.

@@

The results were similar for analyses conducted on Sample 1 and Sample 2
individually. See the online supplemental material for more detail on
these analyses.

@@

@@

The results of this study demonstrate the complex and ambivalent nature
of sexual objectification and additionally highlight the psychological
and social consequences of such objectification processes on women's
social relationships and well-being. We did not find a significant
effect between actor SSO and felt authenticity in the interaction, which
suggests that there is not sufficient evidence to support the claim that
partner objectification is the cause for the diverse range of negative
effects related to interaction inauthenticity. (Move to here: Perhaps
women could be amplifying each others' self-objectification as they get
ready to go out for the night, \emph{before} the theoretical encounters
with men outlined in Gervais et al., 2020 occur.)

>>>>

\abstract{
Empirical evidence has only found links between objectification,
self-objectification, and negative outcomes for women within
interpersonal interactions between male-female pairs. The purpose of the
present study was to extend past research and consider the relationships
between objectification and self-objectification within interactions
between female pairs. Women were brought into the laboratory and
interacted in same-sex dyads. Dyadic analysis was utilized to detect
whether partners' objectification of each other affected state
self-objectification, as well as the resulting feelings of comfort and
authenticity during the interaction. After the interaction, participants
completed a questionnaire measuring, among other things, cognitive
performance, career aspirations, and relationship agency. Results
revealed a significant positive effect of being objectified by a female
partner on women's self-objectification. There was no significant
relationship between self-objectification and authenticity. Further,
although there were significantly negative effects on career aspirations
and relationship agency resulting from a lack of relationship
authenticity, there was no evidence that this is due to feelings of
sexual objectification. The significant partner effect of
objectification on actor self-objectification suggests that women being
objectified by other women may result in feelings of
self-objectification, and such research has powerful implications for
the ways that women interact in both sexual and non-sexual settings.

@@

Roberts and Fredrickson (1997) Objectification Theory, suggests that, in
addition to being steeped in a culture that objectifies women, women are
objectified in actual interpersonal encounters. The negative effects of
this interpersonal objectification are theorized be the strongest when
it is a man doing the objectifying (Gervais, Vescio, \& Allen, 2011),
but what are the consequences when a woman objectifies another woman in
an interaction? Psychological researchers studying the sexual
objectification of women, have in recent years explored the
interpersonal process of the objectification, finding evidence for
self-objectification in the target as a proximal consequence (Garcia,
Earnshaw, \& Quinn, 2016, Strelan and Pagoudis (2018), Riemer, Sáez,
Brock, and Gervais (2020)). For example, in a face-to-face mixed-gender
dyadic study, Garcia et al. (2016) found that men's reported
objectification of their female interaction partner was associated with
increased self-objectification reported by their partner. They did not
include same-gender, woman-woman interacting pairs. Gervais, Sáez,
Riemer, and Klein (2020) have reviewed the literature on interpersonal
self-objectification and has organized the process in a theoretical
model called the Social Interaction Model of Objectification (SIMO), but
this model is currently only relevant for mixed-gender encounters.
Although there is ample evidence that women objectify other women
(Loughnan et al., 2015a; Puvia \& Vaes, 2013), there are currently no
studies investigating the process of interpersonal objectification in
same-gender interactions. Does objectification by a female interaction
partner have the same downstream negative consequences as being
objectified by a male interaction partner? The current dyadic study
addresses this omission using interacting pairs where both members
identify as women.

@@

Harsey and Zurbriggen (2020) found that self-objectification was related
to the objectification of women to a similar degree for men and women.
Additionally, Strelan and Hargreaves (2005) found that the more women
self-objectify, the more they objectify other women. Could the process
work in the other direction as well, that is, could objectification by a
woman produce self-objectification, in a similar manner to mixed-gender
interactions?

Past research has focused on men's objectifying behaviors, with little
attention paid to the role that women may also play in objectifying
other women. Hill and Fischer (2008) assessed women's experiences of
objectification from men independently from women's experiences of
objectification from other women. They found that women may be
socialized not only to see themselves as objects, but perhaps to see
other women as objects as well. This process, however, was not comfirmed
to occur during an actual interpersonal interaction. Loughnan et al.
(2015a) found that women objectify other women to a greater extent than
they objectify men. Women have also been found to objectify
(dehumanized) sexualized targets presented as images (Puvia \& Vaes,
2013), but very little is known about interpersonal sexual
objectification among women and what the proximal effects of this
objectification might be. We do know that objectification has more
adverse consequences for women than men (Gervais et al., 2011; Moradi \&
Huang, 2008; Saguy, Quinn, F Dovidio, \& Pratto, 2010), however, if
women objectify other women, perhaps even to a stronger degree than they
objectify men, this woman to woman objectification might have
detrimental outcomes.

@@

Studies have shown that within social encounters women are gazed at more
than men (Briton \& Hall, 1995), often times feel \enquote{looked at}
within interpersonal interactions (Argyle \& Williams, 1969;
Miles-McLean et al., 2015), and are likely to internalize the
objectifying gaze on their physical self (Puvia \& Vaes, 2013; Young,
1979). Moreover, perhaps the most adverse effect is that being
objectified socializes girls and women to treat themselves as objects to
be looked at and evaluated (Bartky, 1990; Berger, Cohen, \& Zelditch Jr,
1972; B. L. Fredrickson et al., 1998). Gervais et al. (2020) recently
reviewed the research on interpersonal objectification and organized our
current theoretical understanding of this process in the Social
Interaction Model of Objectification (SIMO). What is clear from Gervais
et al. (2020)'s review is that we know little about women-on-women
interpersonal objectification. Could there also be negative consequences
of this type of interpersonal objectification?

Interpersonal objectification can also occur within the context of
romantic relationships, in addition to interactions with strangers.
Among heterosexual male and female college students,
self-objectification has been found to be positively associated with the
extent to which they objectify their romantic partners (Zurbriggen,
Ramsey, \& Jaworski, 2011). Researchers have also studied interpersonal
objectification and self-objectification in romantic relationships
(Strelan \& Pagoudis, 2018). Evidence is mixed, suggesting that women
can sometimes have positive outcomes from appearance valuation within
the context of an established romantic relationship. There is also
evidence that they experience less SSO after appearance comments within
romantic relationships (Meltzer, 2020), but less is known about first
dates. An important caveat to this literature is that it is only about
heterosexual relationships. A recent study has investigated the effect
of self-objectification on the reduction of relationship building skills
in general (including same-sex friendships) (Terán, Jiao, \& Aubrey,
2020). Both romantic relationships and stranger encounters have been
studied in the context of objectification, but neither of these
literatures have investigated specifically same-gender relationships.

There is evidence that the experience of state self-objectification in
mixed-gender contexts (stranger and romantic) has negative consequences
for women, but what about in the context of interactions with other
women? Is the self-objectification experienced in mixed-gender
interactions associated with the same negative process as that
experienced in same-gender interactions? On the one hand, the ample
research demonstrating that the male gaze has a particularly detrimental
effect would suggest no (Gay \& Castano, 2010; Gervais et al., 2011;
Saguy et al., 2010, Calogero (2004); Yilmaz \& Bozo, 2019), but on the
other hand, women do objectify other women (Harsey \& Zurbriggen, 2020;
Loughnan et al., 2015a; Puvia \& Vaes, 2013). When they do it, does it
lead to self-objectification in the same way men's objectification of
women does (Garcia et al., 2016)? If so, does the self-objectification
experienced in these same-gender interactions have the same negative
consequences for authenticity in that interaction?

@@

There is some evidence that reduced authenticity is a consequence of
self-objectification (Garcia et al., 2016; Terán et al., 2020). This
link has been justified by the literature on stigmatized-stigmatizer
interactions (M. R. Hebl \& Dovidio, 2005; Richeson \& Shelton, 2003;
Shelton, Richeson, \& Salvatore, 2005), viewing the experience of being
objectified in an interaction as an identity threat situation
(Miles-McLean et al., 2015; Nadal \& Haynes, 2012). Further, empirical
evidence reveals that objectification manifests through inauthenticity
in romantic relationships (Brunell et al., 2010), adverse attitudes in
regard to career aspirations, and a decrease in concentration and
impairment in female cognitive performance (Kahalon, Shnabel, \& Becker,
2018; D. M. Quinn, Chaudoir, \& Kallen, 2011). When a woman is
objectified by a man, and subsequently experiences self-objectification,
the intergroup nature of this encounter might trigger identity threat.
Thus, perhaps there are fewer negative consequences when a woman is
objectified by another woman, specifically, woman-woman interpersonal
objectification processes might diverge from mix-gender interpersonal
objectification processes precisely because they are not intergroup
interactions (at least with respect to gender identity). A woman
objectified by another woman may not be having the same negative
consequences that cascade from situations that trigger group-based
identity threat (Deaux \& Major, 1987; Dovidio, Hebl, Richeson, \&
Shelton, 2006; M. R. Hebl \& Dovidio, 2005), but feeling like a body,
rather than a full human, in any interaction (intergroup or ingroup) may
be enough to reduce women's feelings of authenticity and social
competence, regardless of the gender of the objectifyer (Terán et al.,
2020; Tolman, Impett, Tracy, \& Michael, 2006).

@@

In the current study, we sought to examine what occurs during an
interaction in which one or both partners are objectifying each other,
similarly to Garcia et al. (2016), but between same-gender interpersonal
interactions among women. Moreover, the current study uses a
face-to-face interaction paradigm and dyadic data analysis techniques to
examine the effects for both women simultaneously. Although the
literature on woman-woman objectification is small, the results are
mixed, and does not cover interpersonal encourters, we expected to
replicate some of the results found in Garcia et al. (2016). Most
importantly, we predicted that being objectified by one's interaction
partner would lead to self-objectification. We also expected that TSO
would moderate this relationship, amplifying the positive assocaition
between being objectified and SSO for women higher in TSO. We
hypothesized that SSO would, in turn, lead to feelings of inauthenticity
(although there is ample support against this prediction from the
identity threat within intergroup interactions literature; Rollero,
2016). Further, we hypothesized that these feelings of inauthencity
would then be associated with reduced feelings of agency in romantic
relationships, reduced career aspiration, and reduced cognitive
performance. In summary, we expected to find a positive relationship
between other-objectification by one's partner and state
self-objectification. We also expected to find a negative relationship
between self-state objectification and interaction authenticity, and
that interaction authenticity will be positively related to cognitive
performance, relationship agency, and career aspirations.

These hypotheses are considered exploratory, given the lack of prior
research theorizing about these linkages.

Here we would like to note that Puvia and Vaes (2013) would
alternatively predict that women's tendency to self-objectify (TSO)
leads them to objectify other women (more precisely, to dehumanize a
sexualized woman). In addition, Strelan and Hargreaves (2005) and Harsey
and Zurbriggen (2020) would predict the TSO to other-objectification
link, further theorizing that this relationships is mediated by the
woman's own state self-objectification (SSO). This limited, but extant,
set of literature on woman-woman objectification points to a possible
alternative model from the model hypothesized in the current study.
Where appropriate, we report results considering the causal directions
implied by this alternative mediation model.

@@

\emph{All} participants were told that they were assigned to the
\enquote{College Relationships} condition. Self-objectification has been
found to occur after a mere relationship prime among women (Sanchez \&
Broccoli, 2008) because, in Westurn culture and beyond, women need to
look attractive to obtain and maintain successful relationship, thus,
the \enquote{College Relationships} condition may heighten
self-objectification and the evaluation of other women in a sexualized
way. The decision was made to ask even heterosexual women to judge their
fellow-woman partners for their potential as romantic partners. We felt
that this prompt would keep the study closest to a replication of the
previous Garcia et al. (2016) version of the study, and strengthen the
potential for objectification that would normally be low in the context
of a psychology laboratory. Recall that, past research has found that
heterosexual women are indeed able to evaluate other women's potential
as romantic partners (i.e., their sexual attractiveness; Puvia \& Vaes,
2013)---indeed, women may be unfortunately quite used to thinking about
their own \enquote{datability,} and we suspect this this habitual
thought pattern will translate to their thoughts about other women.

@@

\subsection{Participants}\label{participants}

Data from two demographically equivalent higher education institutions
in the Northeast United States were combined to create the final
analysis sample (\emph{N =} 64) used in this study. In the measures
section that follows we refer to them as Sample 1 and Sample 2.
Thirty-two previously unacquainted self-identifying female dyads (64
total participants) from institutions in the Northeast of the United
States participated in this study. Sample 1 (\emph{N =} 24) is from a
co-ed liberal arts college and Sample 2 (\emph{N =} 40) is from a
women's liberal arts college. More specifically, twelve of the pairs
were students at a co-ed liberal arts college, while the remaining
twenty pairs attended a women's liberal arts college. {[}Footnote:
Initially, data was collected from same-sex and mixed-sex dyads that
comprised of male and female gendered individuals. Sample 1 originally
consisted of twenty-two pairs, twelve men and thirty-two women.
Twenty-three pairs made up of forty-three women and one man, as well as
two participants who did not identify with either gender category,
formed Sample 2. For consistency, we limited participant data to same
sex female pairs at the two colleges.{]} All results presented below are
from models including sample as a control variable.

The participants were mostly first-year college students, with an
average age of 18.85 (SD = 1.04). The sample was 48.44\% White/European
American, 9.38\% Black/African-American, 28.12\% Asian/Pacific Islander,
9.38\% Latinx, and 4.69\% mixed-race. There were 8 White/White pairs and
4 same race racial minority pairs, for a total of 12 same-race pairs.
The remaining 20 were mixed race pairs, of which 15 were White/racial
minority pairings and 5 were cross-racial minority group pairs. 64.06\%
of the sample identified as heterosexual, and 25\% identified as gay,
lesbian or bisexual.

@@

To measure the participant's objectification of their partner in the
interaction, participants were asked a series of questions about the
frequency of thoughts in relation to multiple characteristics of their
partner (Garcia et al., 2016). Questions included aspects of their
partner's internal traits such as personality, friends, family, and
extracurricular interests, as well as external traits such as body,
appearance, clothing, and body parts. All questions were rated on a
scale from 1 (not at all) to 7 (constantly). Objectification was
measured by taking the difference between the average frequency of
thought about their partner's external traits (\(\alpha\) = 0.79 for
Sample 1, \(\alpha\) = 0.79 for Sample 2) and frequency of thought about
their partner's internal traits (\(\alpha\) = 0.79 for Sample 1,
\(\alpha\) = 0.76 for Sample 2). A positive score in this scale would
indicate that the participant thought about their partner's external
traits more than the partner's internal traits, and a negative score
would indicate the opposite.

@@

\begin{figure}
\includegraphics[width=400px]{theor_figure} \caption{Path Analysis Model from Garcia et al. (2016) study with distinguishable dyads.}\label{fig:theorfigure}
\end{figure}

This study sought to detect whether partners' objectification of one
another affected state self-objectification (SSO). Specifically, we were
interested in testing the relationship between state-other
objectification and SSO, and how SSO in turn, affects feelings of
inauthenticity during the interaction. In addition, we also tested if
the effect of other-objectification in an interaction on SSO is only
present for those women who are high in trait self-objectification.
Further, we investigated the relationships between experiencing
interaction inauthenticity and relationship agency, career aspirations,
and cognitive performance.

While past studies investigating objectification in interactions used
dyadic path analysis (Garcia et al., 2016), the current study used
multilevel modeling proceedures. Dyadic analyses for distinguishable
dyads (e.g., mixed-gender interacting pairs) is more natural in
Structural Equation Modeling (SEM) than it is for indistinguishable
dyads (e.g., same-gender interacting pairs), as we have here (Garcia,
Kenny, \& Ledermann, 2015; Ledermann \& Kenny, 2017). One reason for
this asymmetry is that, due to the arbitrary distinctions made between
\enquote{partner 1} and \enquote{partner 2} in indistinguishable dyads,
many estimates need to be fixed to be equal (i.e., paths, variances,
covariances, endogenous intercepts, and exogenous means), but these
equality constraints should not then be considered in the degrees of
freedom calculations for fit estimations (Olsen \& Kenny, 2006).
Further, Olsen and Kenny (2006) details how a new independence model and
the corresponding fit measure should be re-calculated for
indistinguishable dyads models. The current study uses dyadic multilevel
modeling (MLM) to test all relationships and mediation patterns. The
online supplementary materials contains analysis using SEM. See
Ledermann and Kenny (2017) for a more complete discussion of the
considerations for using SEM versus MLM for dyadic analysis.

Testing our hypotheses in the current, same-gender, sample, involves
using the Actor-Partner Independence Model (APIM) approach for each
outcome variable (i.e., endogenous variable in
Figure~\ref{fig:theorfigure}). Thus, we ran five APIM's to test all the
hypothesized relationships. See Figure~\ref{fig:apim} for a basic APIM
model. The APIM includes effects due to one's own, as well as one's
partner's, predictor variables (\(X\)'s) on the one's own outcome
variable (\(Y\)). The current study deals with indistinguishable dyads,
meaning the designation of who is designated as \enquote{actor} and who
is designated as \enquote{partner} is arbitrary. These analyses are
considered exploratory, given the lack of prior research theorizing
about these linkages.

@@

@@

Objectification among platonic friends: Perhaps women could be
amplifying each others' state self-objectification \emph{before} the
theoretical encounters with men organized in the SIMO occur. In this
way, women's rituals of getting ready together might place even women
with low TSO in a higher SSO group.

@@

The results of this study demonstrate the complex and ambivalent nature
of sexual objectification and additionally highlight the psychological
and social consequences of such objectification processes on women's
social relationships and well-being. We did not find a significant
effect between actor SSO and felt authenticity in the interaction, which
suggests that there is not sufficient evidence to support the claim that
partner objectification is the cause for the diverse range of negative
effects related to interaction inauthenticity.
